Macro Impacts

### 2016-08-03 Macro Impacts Revised.tex
Attenuation *possibly* figured out

To keep track of the quality factor, we need to be able to pointwise multiply $e^{-\f{k}{2Q}t}$ by $A(k)$. The problem is that Matlab doesn't really do functions... Well... Suppose we start with an amplitude $A(0,k)$, and that the ray propagates through a region of quality factor $Q_1$ for time $\Delta t_1$. Then the amplitude after the propagation is $A(\Delta t_1, k)=A(0,k)F_1 e^{-\f{k}{2Q_1}\Delta t_1}$, where $F_1$ is a transmission or reflection coefficient, which is not a function of $k$. Now, suppose it propagates through another region with quality factor $Q_2$ for $\Delta t_2$. Then we have $A(\Delta t_1 + \Delta t_2, k)=A(0,k)F_1F_2 e^{-\f{k}{2}\p{\f{\Delta t_1}{Q_1}+\f{\Delta t_2}{Q_2}}}$. That's straight forward enough. Therefore, to understand what happens to a general ray's amplitude, we only need to know two things:

@@

\f{T}{Q} &= \sum_i\f{\Delta t_i}{Q_i}\,.

@@

p_0(m)^2 = \f{ K_m^2\rho_m}{\lambda_m + 2\mu_m}\f{2l^2\sigma_Xv_X^2}{\int r\diff r f(r)^2}\,.

@@

\notag &=\f{Q}{NT^2}F_{m,n}^2e^{-\f{k_0^2T}{Q}}\ps{4T(1-\cos(l k_0)) + \I e^{\f{k_0 T}{Q}} l Q\p{2\te{Ei}\ps{-\I l \notag k_0-\f{k_0 T}{Q}} - 2\te{Ei}\ps{\I l k_0 - \f{k_0 T}{Q}} + \right.\right.\\&\left.\left.+\ln\ps{-\f{Q}{\I l Q+T}}+\ln\notag \ps{-\f{\I l Q + T}{Q}} + 2\ln\ps{\I l - \f{T}{Q}}} - 4 l(Q + k_0 T)\te{Si}\ps{l k_0}}\times\\
&\times \underbrace{\f{(b^2f_1(b)-a^2f_1(a))^2}{\int r\diff r f(r)^2}}_{A}\rho_m2l^2\sigma_Xv_X^2\,.

@@

E_m'\leq F_{m,n}^24 l p_0^2 \f{Q^2}{T^2} \arctan\ps{l\f{Q}{T}} A\rho_m2l^2\sigma_Xv_X^2\,.

@@

Now, for a ray that immediately hits the surface, its energy must be the same as when it started. This corresponds to $T = \to 0$, and

@@

E_m = 4l E_mAF_{m,n}^2\p{-2k_0+k_0\cos(h k_0)+\f{\sin(h k_0)}{h}+h k_0^2\te{Si}(h k_0)}

@@

where we've taken $l\rho_m\sigma_Xv_X^2=E_m$.

@@

 \p{4lF_{m,n}^2\p{-2k_0+k_0\cos(h k_0)+\f{\sin(h k_0)}{h}+h k_0^2\te{Si}(h k_0)}}^{-1}=  A

@@

These units are way wrong. Gonna fix tomorrow.

@@

The model above produces a set of positions on a sphere along with their associated times of impact and relative amplitudes at particular frequencies. The detectable energy from any given ray is the energy that is deposited in the frequencies detectable to a seismometer. According to Wikipedia, some seismometers can measure frequencies from 500 Hz to 0.00118 Hz. The lunar seismometers are considerably less sensitive, and have useful range 0.004 Hz to 2 Hz according to Wikipedia. There seems to be a total of 12 lunar seismometers.

>>>>

To keep track of the quality factor, we need to be able to pointwise multiply $e^{-\f{vk}{2Q}t}$ by $A(k)$. The problem is that Matlab doesn't really do functions... Well... Suppose we start with an amplitude $A(0,k)$, and that the ray propagates through a region of quality factor $Q_1$ for time $\Delta t_1$. Then the amplitude after the propagation is $A(\Delta t_1, k)=A(0,k)F_1 e^{-\f{kv}{2Q_1}\Delta t_1}$, where $F_1$ is a transmission or reflection coefficient, which is not a function of $k$. Now, suppose it propagates through another region with quality factor $Q_2$ for $\Delta t_2$. Then we have $A(\Delta t_1 + \Delta t_2, k)=A(0,k)F_1F_2 e^{-\f{k}{2}\p{\f{\Delta t_1v_1}{Q_1}+\f{\Delta t_2v_2}{Q_2}}}$. We take $v_i$ to be the average velocity in a given region. This is a fine approximation since the velocity doesn't vary much within the moon. That's straight forward enough. Therefore, to understand what happens to a general ray's amplitude, we only need to know two things:

@@

\f{T}{Q} &= \sum_i\f{\Delta t_i v_i}{Q_i}\,.

@@

p_0(m)^2 = \f{ K_m^2}{\lambda_m + 2\mu_m}\f{2\rho_m\sigma_Xv_X^2}{\int r\diff r f(r)^2}\,.

@@

\notag &=\f{Q}{NT^2}e^{-\f{k_0^2T}{Q}}\ps{4T(1-\cos(l k_0)) + \I e^{\f{k_0 T}{Q}} l Q\p{2\te{Ei}\ps{-\I l \notag k_0-\f{k_0 T}{Q}} - 2\te{Ei}\ps{\I l k_0 - \f{k_0 T}{Q}} + \right.\right.\\&\left.\left.+\ln\ps{-\f{Q}{\I l Q+T}}+\ln\notag \ps{-\f{\I l Q + T}{Q}} + 2\ln\ps{\I l - \f{T}{Q}}} - 4 l(Q + k_0 T)\te{Si}\ps{l k_0}}\times\\
&\times (b^2f_1(b)-a^2f_1(a))^2\f{\lambda_m+2\mu_m}{K_m^2}F_{m,n}^2p_0(m)^2\,,\\
\notag &=\f{Q}{NT^2}e^{-\f{k_0^2T}{Q}}\ps{4T(1-\cos(l k_0)) + \I e^{\f{k_0 T}{Q}} l Q\p{2\te{Ei}\ps{-\I l \notag k_0-\f{k_0 T}{Q}} - 2\te{Ei}\ps{\I l k_0 - \f{k_0 T}{Q}} + \right.\right.\\&\left.\left.+\ln\ps{-\f{Q}{\I l Q+T}}+\ln\notag \ps{-\f{\I l Q + T}{Q}} + 2\ln\ps{\I l - \f{T}{Q}}} - 4 l(Q + k_0 T)\te{Si}\ps{l k_0}}\times\\
&\times \underbrace{\f{(b^2f_1(b)-a^2f_1(a))^2}{\int r\diff r f(r)^2}}_{A}F_{m,n}^2 2\rho_m\sigma_Xv_X^2\,.

@@

E_m' =  4 l  \f{Q^2}{T^2} \arctan\ps{l\f{Q}{T}} F_{m,n}^2 A2\rho_m\sigma_Xv_X^2\,.

@@

How do we remove $A$ from this equation?
Now, for a ray that immediately hits the surface, its energy must be the same as when it started. This corresponds to $T = \to 0$ and $F_{m,n} = 1$

@@

E_m = 4\p{-2k_0+k_0\cos(l k_0)+\f{\sin(h k_0)}{l}+l k_0^2\te{Si}(l k_0)} A \td{E}{x}{}(m)\,.

@@

@@

A = \f{l}{4} \p{-2k_0+k_0\cos(l k_0)+\f{\sin(h k_0)}{l}+l k_0^2\te{Si}(l k_0)}^{-1}\,.

@@

This is, really, an upper bound to what $A$ can be, since $A$ really depends on $m$. When we fix $A$, we lose information about how the frequencies are distributed, and we, in some sense, assume that all the energy from the explosion is initially concentrated in the lower frequencies. This is good for an upper estimate of the energy felt by seismometers. Note, for bigger $k_0$, $E_m$ experiences a faster drop-off as $T$ gets big.

@@

The model above produces a set of positions on a sphere along with their associated times of impact and relative amplitudes at particular frequencies. The detectable energy from any given ray is the energy that is depositied in the frequencies detectable to a seismometer. According to Wikipedia, some seismometers can measure frequencies from 500 Hz to 0.00118 Hz. The lunar seismometers are considerably less sensitive, and have useful range 0.004 Hz to 2 Hz according to Wikipedia. There seems to be a total of 12 lunar seismometers.
